---
title: "p8105_homework1_xl2834"
author: "Xiaoyu Li"
date: "9/15/2020"
output: github_document
---


```{r}
library(tidyverse)
```

## Problem 1

Create a data frame with the specified elements.
```{r}
prob1_df = 
  tibble(
    samp = rnorm(10),
    vec_logical = samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j" ),
    vec_factor = factor(c("low", "mod", "low", "mod", "high", "low",                           "mod", "high", "high", "low"))
)
```


Take the mean of the variables in the data frame.
```{r}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, vec_logical))
mean(pull(prob1_df, vec_char))
mean(pull(prob1_df, vec_factor))
```

I can take the mean of numeric sample and logical vector, but not the character and factor.

Convert the logical vector to other types.
```{r, results='hide', warning=FALSE}
as.numeric(pull(prob1_df,vec_logical)) * pull(prob1_df, samp)
as.factor(pull(prob1_df,vec_logical)) * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df,vec_logical))) * pull(prob1_df, samp)
```

I can multiply the numeric sample with the logical vector after it was converted to numeric or converted to factor and then numeric, but not after it was converted to a factor. 

This can help to explain why I cannot take the mean of a factor (because the operators like "+" and "*" are not meaningful for factors).


## Problem 2

Download the package containing the penguins dataset.
```{r}
data("penguins", package = "palmerpenguins")
```

This dataset include variables `r ls(penguins)`. 
The species of penguins include `r unique (penguins$species)`. The islands include `r unique (penguins$island)`. The maximum value of bill length is `r max(penguins$bill_length_mm,na.rm = TRUE)`, and the minimum value is `r min(penguins$bill_length_mm,na.rm = TRUE)`.
The size of the dataset is `r nrow(penguins)` * `r ncol(penguins)`.
The mean flipper length is `r mean(penguins$flipper_length_mm,na.rm = TRUE)`.

#### Plot

Make a scatterplot.
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

```

Save the plot.
```{r}
ggsave("scatterplot.pdf")
```

